### 2022-03-29 manuscript.tex
Update 02.tubaseq.md [ci skip]
This build is based on
https://github.com/scherzerthesis/thesis/commit/a6a894e99cfd62a104f67d3ecf75754b618afe0c.

This commit was created by the following CI build and job:
https://github.com/scherzerthesis/thesis/commit/a6a894e99cfd62a104f67d3ecf75754b618afe0c/checks
https://github.com/scherzerthesis/thesis/actions/runs/2055526507

This manuscript
(\href{https://scherzerthesis.github.io/thesis/v/c1eb46dca3a6035d359424450c0f133449d977a6/}{permalink})
was automatically generated
from \href{https://github.com/scherzerthesis/thesis/tree/c1eb46dca3a6035d359424450c0f133449d977a6}{scherzerthesis/thesis@c1eb46d}
on March 28, 2022.

@@

Recently, Monte Winslow and his lab at Stanford University has developed \textbf{Tu}mor- \textbf{Ba}rcode \textbf{Seq}uencing (TUBA-SEQ), a rapid, multiplexed, quantitative assay for analyzing the contribution of tumor-suppressor loss \emph{in vivo}. This technique works by delivering a uniquely barcoded lentiviral endoing Cre recombinase, to activate Cre-mediated oncogenes, and a guide RNA, to inactivate a single tumor suppressor gene. Therefore, each tumor will be driven by the same oncogene, but will have different tumor suppressors inactivated. After a period of time (6-24 weeks) tissue is harvested for whole-lung DNA extraction and targeted DNA-sequencing is performed to quantify relative tumor size to spike-in controls(Rogers \emph{et al.}, 2017). Therefore, contribution of a single tumor suppressor can be quantified in single animal.

@@

Gross histologic images of experimental lungs are show an increase in tumor burden with both time (Figure xxx) and CAS9 status (figure xxx). Therefore, based on macroscopic inspection, at least one tumor suppressor appeared to cooperate with BRAF \textsuperscript{V600E} to increase tumor burden. In order to determine the fitness landscape of any one particular tumor suppressor, we calculated the relative number of tumor cells in a given tumor using the TUBA-SEQ genomics pipeline. After 8 weeks post tumor initiation there no statistically significant contribution of tumor suppressor loss (Figure). SETD2 loss was trending towards statistical significance. However, 16 weeks post initiation revealed both SETD2 loss and APC loss cooperated with BRAF \textsuperscript{V600E} to form larger tumors (Figure x). APC loss was a suspected hit as we have previously shown that activation of Wnt-beta-catenin signaling promoted BRAF\textsuperscript{V600E} tumorigenesis. As APC loss is predicted to increase the activity of Beta-catenin signaling, this result was unsurprising to us. However, loss of SETD2 was a novel finding in the context of BRAF \textsuperscript{V600E} lung tumorigenesis.

@@

The mutational landscape of lung adenocarcinoma is complex due lifelong bombardment on DNA by carcinogen ingested into the distal lung epithelium{[}23770567{]}. While understanding which altered genes play causal roles in tumor formation and progression is required for personalized medicine, the noisy mutation profiles in lung adenocarcinoma makes it hard to do so. However, it is clear that there is a central role for the MAPK pathway during the initiation and progression of lung adenocarcinoma (Cicchini \emph{et al.}, 2017). Oncogenic BRAF\textsuperscript{V600E} is a major driver of lung adenocarcinoma. However, mouse models with activation of BRAF \textsuperscript{V600E} only lead to benign tumorigenesis(Shai \emph{et al.}, 2015; Dankort \emph{et al.}, 2007). Although it is known that activation of WNT and PI3k signaling, as well as lost of \emph{Trp53} and \emph{Cdkn2a} genes allow BRAF\textsuperscript{V600E} benign lung tumors to progress to malignant disease, it is a time-consuming and laborious process to effects of individual tummor suppressors in GEMMs.Therefore, we sought out to systematically test a variety of known tumor suppressors in lung adenocarcinoma. Furthermore,untilnow, we did not know the absolute number of cells in a BRAF\textsuperscript{V600E} adenoma, only its area, which is confounded by tissue sectioning as well as stromal cell contamination. Prior to TUBA-SEQ, it was very difficult to accurately determine the size of BRAF\textsuperscript{V600E} adenomas. Based on population doublings, as well as cross sections of BRAF \textsuperscript{V600E} tumors in histological sections, it is estimated that BRAF\textsuperscript{V600E}-driven lung adenomas grow about 15-20 populations doublings and reaching a size of 1mm\textsuperscript{3}(Juan \emph{et al.}, 2014). With Tuba-Seq, we accurately discerned the absolute size of all BRAF\textsuperscript{V600E} lung adenomas, regardless of size. Interestingly, our results have revealed BRAF\textsuperscript{V600E}-driven tumors widely differ in absloute tumor cell number. Our results support the hypothesis that each tumor-initiating cell has its own unique barriers for tumor outgrowth. This heterogenous reponse to BRAF\textsuperscript{V600E} activation likely reflects intrinisic differences between individual cell states. Indeed, tumorigenic potential can vary greatly due to different cell types (AT2 vs Club cell vs AT1), or can be influenced by subtle differences within a cell type, such proximity to supporting stromal cells such as fibroblast or immune cells. Since we used a ubiquitous promoter to drive BRAF\textsuperscript{V600E} and CAS9 expression, it is likely that not all sequence reads come from adenomas or tumors, but rather from normal cell turnover during tissue homeostasis. Therefore, in out quantifcations and analysis, we focused on tumors that were greater than 5,000 cells.

>>>>

This manuscript
(\href{https://scherzerthesis.github.io/thesis/v/a6a894e99cfd62a104f67d3ecf75754b618afe0c/}{permalink})
was automatically generated
from \href{https://github.com/scherzerthesis/thesis/tree/a6a894e99cfd62a104f67d3ecf75754b618afe0c}{scherzerthesis/thesis@a6a894e}
on March 28, 2022.

@@

Recently, Monte Winslow and his lab at Stanford University has developed \textbf{Tu}mor- \textbf{Ba}rcode \textbf{Seq}uencing (Tuba-Seq), a rapid, multiplexed, quantitative assay for analyzing the contribution of tumor-suppressor loss \emph{in vivo}. This technique works by delivering a uniquely barcoded lentiviral endoing Cre recombinase, to activate Cre-mediated oncogenes, and a guide RNA, to inactivate a single tumor suppressor gene. Therefore, each tumor will be driven by the same oncogene, but will have different tumor suppressors inactivated. After a period of time (6-24 weeks) tissue is harvested for whole-lung DNA extraction and targeted DNA-sequencing is performed to quantify relative tumor size to spike-in controls(Rogers \emph{et al.}, 2017). Therefore, contribution of a single tumor suppressor can be quantified in single animal.

@@

Gross histologic images of experimental lungs are show an increase in tumor burden with both time (Figure xxx) and CAS9 status (figure xxx). Therefore, based on macroscopic inspection, at least one tumor suppressor appeared to cooperate with BRAF \textsuperscript{V600E} to increase tumor burden. In order to determine the fitness landscape of any one particular tumor suppressor, we calculated the relative number of tumor cells in a given tumor using the Tuba-Seq genomics pipeline. After 8 weeks post tumor initiation there no statistically significant contribution of tumor suppressor loss (Figure). SETD2 loss was trending towards statistical significance. However, 16 weeks post initiation revealed both SETD2 loss and APC loss cooperated with BRAF \textsuperscript{V600E} to form larger tumors (Figure x). APC loss was a suspected hit as we have previously shown that activation of Wnt-beta-catenin signaling promoted BRAF\textsuperscript{V600E} tumorigenesis. As APC loss is predicted to increase the activity of Beta-catenin signaling, this result was unsurprising to us. However, loss of SETD2 was a novel finding in the context of BRAF \textsuperscript{V600E} lung tumorigenesis.

@@

The mutational landscape of lung adenocarcinoma is complex due lifelong bombardment on DNA by carcinogen ingested into the distal lung epithelium{[}23770567{]}. While understanding which altered genes play causal roles in tumor formation and progression is required for personalized medicine, the noisy mutation profiles in lung adenocarcinoma makes it hard to do so. However, it is clear that there is a central role for the MAPK pathway during the initiation and progression of lung adenocarcinoma (Cicchini \emph{et al.}, 2017). Oncogenic BRAF\textsuperscript{V600E} is a major driver of lung adenocarcinoma. However, mouse models with activation of BRAF \textsuperscript{V600E} only lead to benign tumorigenesis(Shai \emph{et al.}, 2015; Dankort \emph{et al.}, 2007). Although it is known that activation of WNT and PI3k signaling, as well as lost of \emph{Trp53} and \emph{Cdkn2a} genes allow BRAF\textsuperscript{V600E} benign lung tumors to progress to malignant disease, it is a time-consuming and laborious process to effects of individual tummor suppressors in GEMMs.Therefore, we sought out to systematically test a variety of known tumor suppressors in lung adenocarcinoma. Furthermore,untilnow, we did not know the absolute number of cells in a BRAF\textsuperscript{V600E} adenoma, only its area, which is confounded by tissue sectioning as well as stromal cell contamination. Prior to Tuba-Seq, it was very difficult to accurately determine the size of BRAF\textsuperscript{V600E} adenomas. Based on population doublings, as well as cross sections of BRAF \textsuperscript{V600E} tumors in histological sections, it is estimated that BRAF\textsuperscript{V600E}-driven lung adenomas grow about 15-20 populations doublings and reaching a size of 1mm\textsuperscript{3}(Juan \emph{et al.}, 2014). With Tuba-Seq, we accurately discerned the absolute size of all BRAF\textsuperscript{V600E} lung adenomas, regardless of size. Interestingly, our results have revealed BRAF\textsuperscript{V600E}-driven tumors widely differ in absloute tumor cell number. Our results support the hypothesis that each tumor-initiating cell has its own unique barriers for tumor outgrowth. This heterogenous reponse to BRAF\textsuperscript{V600E} activation likely reflects intrinisic differences between individual cell states. Indeed, tumorigenic potential can vary greatly due to different cell types (AT2 vs Club cell vs AT1), or can be influenced by subtle differences within a cell type, such proximity to supporting stromal cells such as fibroblast or immune cells. Since we used a ubiquitous promoter to drive BRAF\textsuperscript{V600E} and CAS9 expression, it is likely that not all sequence reads come from adenomas or tumors, but rather from normal cell turnover during tissue homeostasis. Therefore, in out quantifcations and analysis, we focused on tumors that were greater than 5,000 cells.
